---
title: "Impute hormones by cycle day"
author: "Ruben Arslan"
date: "1/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
```

```{r}
library(tidyverse)
library(codebook)
library(brms)

plot_curve <- function(model, method = "fitted") {
  cond <- tibble(.rows = 1)
  cond[[names(model$data)[2]]] <- "none"
  plot(conditional_effects(model, method = method,
                conditions = cond, 
                spaghetti = TRUE, nsamples = 100), 
              points = TRUE, 
              point_args = list(alpha = 0.05),
              spaghetti_args = list(alpha = 0.01, colour = '#FF000007'), 
              line_args = list(size = 0), 
              plot = FALSE)[[1]]
}
rmse <- function(error) { sqrt(mean(error^2, na.rm = TRUE)) }
rmse_brms <- function(model) {
  resid <- residuals(model, summary = F)
  rmse_samples <- apply(resid, 1, rmse)
  rmse_ci <- rstantools::posterior_interval(as.matrix(rmse_samples), prob = 0.95)
  sprintf("%.2f [%.2f;%.2f]", mean(rmse_samples), rmse_ci[1], rmse_ci[2]) 
}
mae <- function(error) { mean(abs(error), na.rm = TRUE) }
mae_brms <- function(model) {
  resid <- residuals(model, summary = F)
  mae_samples <- apply(resid, 1, mae)
  mae_ci <- rstantools::posterior_interval(as.matrix(mae_samples), prob = 0.95)
  sprintf("%.2f [%.2f;%.2f]", mean(mae_samples), mae_ci[1], mae_ci[2]) 
}

loo_ci <- function(model) {
  loo_R <- loo_R2(model, re_formula = NA)
  loo_R <- sqrt(loo_R)
  sprintf("%.2f [%.2f;%.2f]", loo_R[,"Estimate"], loo_R[,"Q2.5"], loo_R[,"Q97.5"]) 
}
```


```{r}
biocycle <- readRDS("biocycle.rds")
options(mc.cores = parallel::detectCores(), brms.backend = "cmdstanr", brms.file_refit = "on_change")
rstan::rstan_options(auto_write = TRUE)

biocycle <- biocycle %>% filter(between(cycle_length, 20, 35))
biocycle$logE_P <- log(biocycle$E_P)
biocycle$logE <- log(biocycle$total_estradiol)
biocycle$logP <- log(biocycle$progesterone)
biocycle$logT <- log(biocycle$Testo)
biocycle$logFE <- log(biocycle$estradiol)
biocycle$logFT <- log(biocycle$`Free T (ng/dL)`)
biocycle$fertile_lh <- NULL
```

## Days
```{r}
bc_days <- data.frame(
	bc_day = c(-28:-1, -29:-40),
	prc_stirn_bc = c(.01, .01, .02, .03, .05, .09, .16, .27, .38, .48, .56, .58, .55, .48, .38, .28, .20, .14, .10, .07, .06, .04, .03, .02, .01, .01, .01, .01, rep(.01, times = 12)),
# 	                rep(.01, times = 70)), # gangestad uses .01 here, but I think such cases are better thrown than kept, since we might simply have missed a mens
	prc_wcx_bc = c(.000, .000, .001, .002, .004, .009, .018, .032, .050, .069, .085, .094, .093, .085, .073, .059, .047, .036, .028, .021, .016, .013, .010, .008, .007, .006, .005, .005, rep(.005, times = 12))
) %>% arrange(bc_day)

fc_days <- data.frame(
	fc_day = c(0:39),
	prc_stirn_fc = c(.01, .01, .02, .03, .05, .09, .16, .27, .38, .48, .56, .58, .55, .48, .38, .28, .20, .14, .10, .07, .06, .04, .03, .02, .01, .01, .01, .01, rep(.01, times = 12)),
# 	                rep(.01, times = 70)), # gangestad uses .01 here, but I think such cases are better thrown than kept, since we might simply have missed a mens
	prc_wcx_fc = c(.000, .000, .001, .002, .004, .009, .018, .032, .050, .069, .085, .094, .093, .085, .073, .059, .047, .036, .028, .021, .016, .013, .010, .008, .007, .006, .005, .005, rep(.005, times = 12))
)

lh_days <- tibble(
    conception_risk_lh = c(rep(0,8), 0.00, 0.01, 0.02, 0.06, 0.16, 0.20, 0.25, 0.24, 0.10, 0.02, 0.02, rep(0,12)),
    lh_day = -15:15
  ) %>% 
    mutate(fertile_lh = conception_risk_lh/max(conception_risk_lh)*0.25/0.31)
biocycle <- biocycle %>% left_join(lh_days %>% select(lh_day, fertile_lh), by = "lh_day")
```


## Impute E
### Backward counting
```{r}
(mod_e_bc = brm(bf(logE | cens(estradiol_cens) ~ s(bc_day) + (1 | id) + (1|id:cycle)),  
                data = biocycle, control = list(adapt_delta = 0.95, max_treedepth = 20),
                file ='models/mod_e_bc'))

performance::variance_decomposition(mod_e_bc)
# performance::model_performance(mod_e_bc, metrics = c("RMSE", "R2", "R2_adj"))
loo_R2(mod_e_bc, re_formula = NA)
rmse_brms(mod_e_bc)
mae_brms(mod_e_bc)
plot_curve(mod_e_bc)

bc_days[,paste0("est_estradiol_bc", c("", "_se", "_low", "_high"))] <- round(predict(mod_e_bc, newdata = bc_days, re_formula = NA),2)
```

### Forward counting
```{r}
(mod_e_fc = brm(bf(logE | cens(estradiol_cens) ~ s(fc_day) + (1 | id) + (1|id:cycle)),  
                data = biocycle, control = list(adapt_delta = 0.95, max_treedepth = 20),
                file ='models/mod_e_fc'))

loo_R2(mod_e_fc, re_formula = NA)
rmse_brms(mod_e_fc)
mae_brms(mod_e_fc)
plot_curve(mod_e_fc, method = "fitted")

fc_days[,paste0("est_estradiol_fc", c("", "_se", "_low", "_high"))] <- round(predict(mod_e_fc, newdata = fc_days, re_formula = NA),2)
```

### Relative to LH surge
```{r}
(mod_e_lh = brm(bf(logE | cens(estradiol_cens) ~ s(lh_day) + (1 | id) + (1|id:cycle)),  
                data = biocycle %>% filter(between(lh_day, -20, 14)), control = list(adapt_delta = 0.95, max_treedepth = 20),
                file ='models/mod_e_lh')
)

loo_R2(mod_e_lh, re_formula = NA)
rmse_brms(mod_e_lh)
mae_brms(mod_e_lh)
plot_curve(mod_e_lh, method = "fitted")

lh_days[,paste0("est_estradiol_lh", c("", "_se", "_low", "_high"))] <- round(predict(mod_e_lh, newdata = lh_days, re_formula = NA),2)
```

## Impute FE
### Backward counting
```{r}
(mod_fe_bc = brm(bf(logFE | cens(estradiol_cens) ~ s(bc_day) + (1 | id) + (1|id:cycle)),  
                data = biocycle, control = list(adapt_delta = 0.95, max_treedepth = 20),
                file ='models/mod_fe_bc'))

loo_R2(mod_fe_bc, re_formula = NA)
rmse_brms(mod_fe_bc)
mae_brms(mod_fe_bc)
plot_curve(mod_fe_bc, method = "fitted")

bc_days[,paste0("est_free_estradiol_bc", c("", "_se", "_low", "_high"))] <- round(predict(mod_fe_bc, newdata = bc_days, re_formula = NA),2)
```

### Forward counting
```{r}
mod_fe_fc = brm(bf(logFE | cens(estradiol_cens) ~ s(fc_day) + (1 | id) + (1|id:cycle)),  
                data = biocycle, control = list(adapt_delta = 0.95, max_treedepth = 20),
                file ='models/mod_fe_fc')

loo_R2(mod_fe_fc, re_formula = NA)
rmse_brms(mod_fe_fc)
mae_brms(mod_fe_fc)
plot_curve(mod_fe_fc, method = "fitted")

fc_days[,paste0("est_free_estradiol_fc", c("", "_se", "_low", "_high"))] <- round(predict(mod_fe_fc, newdata = fc_days, re_formula = NA),2)
```

### Relative to LH surge
```{r}
mod_fe_lh = brm(bf(logFE | cens(estradiol_cens) ~ s(lh_day) + (1 | id) + (1|id:cycle)),  
                data = biocycle %>% filter(between(lh_day, -20, 14)), control = list(adapt_delta = 0.95, max_treedepth = 20),
                file ='models/mod_fe_lh')

loo_R2(mod_fe_lh, re_formula = NA)
rmse_brms(mod_fe_lh)
mae_brms(mod_fe_lh)
plot_curve(mod_fe_lh, method = "fitted")

lh_days[,paste0("est_free_estradiol_lh", c("", "_se", "_low", "_high"))] <- round(predict(mod_fe_lh, newdata = lh_days, re_formula = NA),2)
```

## Impute P
### Backward counting
```{r}
mod_p_bc = brm(bf(logP | cens(progesterone_cens) ~ s(bc_day) + (1 | id) + (1|id:cycle)),  
                data = biocycle, control = list(adapt_delta = 0.95, max_treedepth = 20),
                file ='models/mod_p_bc')
plot_curve(mod_p_bc, method = "fitted")

bc_days[,paste0("est_progesterone_bc", c("", "_se", "_low", "_high"))] <- round(predict(mod_p_bc, newdata = bc_days, re_formula = NA),2)
```

### Forward counting
```{r}
mod_p_fc = brm(bf(logP | cens(progesterone_cens)  ~ s(fc_day) + (1 | id) + (1|id:cycle)),  
                data = biocycle, control = list(adapt_delta = 0.95, max_treedepth = 20),
                file ='models/mod_p_fc')
plot_curve(mod_p_fc, method = "fitted")

fc_days[,paste0("est_progesterone_fc", c("", "_se", "_low", "_high"))] <- round(predict(mod_p_fc, newdata = fc_days, re_formula = NA),2)
```


### Relative to LH surge
```{r}
mod_p_lh = brm(bf(logP | cens(progesterone_cens)  ~ s(lh_day) + (1 | id) + (1|id:cycle)),  
                data = biocycle %>% filter(between(lh_day, -20, 14)), control = list(adapt_delta = 0.95, max_treedepth = 20),
                file ='models/mod_p_lh')

plot_curve(mod_p_lh, method = "fitted")

lh_days[,paste0("est_progesterone_lh", c("", "_se", "_low", "_high"))] <- round(predict(mod_p_lh, newdata = lh_days, re_formula = NA),2)
```

## New PBFW (probability of being in fertile window)
### Backward counting
```{r}
mod_pbfw_bc = brm(bf(fertile_lh ~ s(bc_day) + (1 | id) + (1|id:cycle)),
                data = biocycle, 
                file ='models/mod_pbfw_bc')
plot_curve(mod_pbfw_bc, method = "fitted")

bc_days[,paste0("est_pbfw_bc", c("", "_se", "_low", "_high"))] <- round(predict(mod_pbfw_bc, newdata = bc_days, re_formula = NA),2)
```

### Forward counting
```{r}
mod_pbfw_fc = brm(bf(fertile_lh ~ s(fc_day)  + (1 | id) + (1|id:cycle)),
                data = biocycle, 
                file ='models/mod_pbfw_fc')
plot_curve(mod_pbfw_fc, method = "fitted")

fc_days[,paste0("est_pbfw_fc", c("", "_se", "_low", "_high"))] <- round(predict(mod_pbfw_fc, newdata = fc_days, re_formula = NA),2)
```

### Relative to LH surge
Not done, because predictor = criterion.



## Impute T
### Backward counting
```{r}
mod_t_bc = brm(bf(logT | cens(Testo_cens)  ~ s(bc_day) + (1 | id) + (1|id:cycle)),  
                data = biocycle, control = list(adapt_delta = 0.95, max_treedepth = 20),
                file ='models/mod_t_bc')

loo_R2(mod_t_bc, re_formula = NA)
rmse_brms(mod_t_bc)
mae_brms(mod_t_bc)
plot_curve(mod_t_bc, method = "fitted")


bc_days[,paste0("est_testosterone_bc", c("", "_se", "_low", "_high"))] <- round(predict(mod_t_bc, newdata = bc_days, re_formula = NA),2)
```

### Forward counting
```{r}
mod_t_fc = brm(bf(logT | cens(Testo_cens)  ~ s(fc_day) + (1 | id) + (1|id:cycle)),  
                data = biocycle, control = list(adapt_delta = 0.95, max_treedepth = 20),
                file ='models/mod_t_fc')

plot_curve(mod_t_fc, method = "fitted")

fc_days[,paste0("est_testosterone_fc", c("", "_se", "_low", "_high"))] <- round(predict(mod_t_fc, newdata = fc_days, re_formula = NA),2)
```

### Relative to LH surge
```{r}
mod_t_lh = brm(bf(logT | cens(Testo_cens)  ~ s(lh_day) + (1 | id) + (1|id:cycle)),  
                data = biocycle %>% filter(between(lh_day, -20, 14)), control = list(adapt_delta = 0.95, max_treedepth = 20),
                file ='models/mod_t_lh')

plot_curve(mod_t_lh, method = "fitted")

lh_days[,paste0("est_testosterone_lh", c("", "_se", "_low", "_high"))] <- round(predict(mod_t_lh, newdata = lh_days, re_formula = NA),2)
```

## Impute FE
### Backward counting
```{r}
mod_ft_bc = brm(bf(logFT | cens(Testo_cens)  ~ s(bc_day) + (1 | id) + (1|id:cycle)),  
                data = biocycle, control = list(adapt_delta = 0.95, max_treedepth = 20),
                file ='models/mod_ft_bc')

plot_curve(mod_ft_bc, method = "fitted")

bc_days[,paste0("est_free_testosterone_bc", c("", "_se", "_low", "_high"))] <- round(predict(mod_ft_bc, newdata = bc_days, re_formula = NA),2)
```

### Forward counting
```{r}
mod_ft_fc = brm(bf(logFT | cens(Testo_cens)  ~ s(fc_day) + (1 | id) + (1|id:cycle)),  
                data = biocycle, control = list(adapt_delta = 0.95, max_treedepth = 20),
                file ='models/mod_ft_fc')

plot_curve(mod_ft_fc, method = "fitted")

fc_days[,paste0("est_free_testosterone_fc", c("", "_se", "_low", "_high"))] <- round(predict(mod_ft_fc, newdata = fc_days, re_formula = NA),2)
```

### Relative to LH surge
```{r}
mod_ft_lh = brm(bf(logFT | cens(Testo_cens)  ~ s(lh_day) + (1 | id) + (1|id:cycle)),  
                data = biocycle %>% filter(between(lh_day, -20, 14)), control = list(adapt_delta = 0.95, max_treedepth = 20),
                file ='models/mod_ft_lh')

plot_curve(mod_ft_lh, method = "fitted")

lh_days[,paste0("est_free_testosterone_lh", c("", "_se", "_low", "_high"))] <- round(predict(mod_ft_lh, newdata = lh_days, re_formula = NA),2)
```


## Label variables
```{r}
lh_days <- as.data.frame(lh_days)
var_label(fc_days) <- list(fc_day = "Forward-counted cycle day. Counted from last (recalled) menstrual onset, starting at zero.",
                           prc_stirn_fc = "Probability of being in the fertile window determined according to the forward-counted cycle days, probabilities from Gangestad et al. 2016 based on Stirnemann et al. 2013",
                           prc_wcx_fc = "Probability of being in the fertile window determined according to the forward-counted cycle days, probabilities from Gangestad et al. 2016 based on Wilcox et al. 2001",
                           est_estradiol_fc = paste("Imputed serum estradiol level based on the BioCycle study, loo-R =", loo_ci(mod_e_fc)),
                           est_estradiol_fc_se = "Standard error: Imputed serum estradiol.",
                           est_estradiol_fc_low = "Lower 95% CI: Imputed serum estradiol.",
                           est_estradiol_fc_high = "Upper 95% CI: Imputed serum estradiol.",
                           est_free_estradiol_fc = paste("Imputed free serum estradiol level based on the BioCycle study, loo-R =", loo_ci(mod_fe_fc)),
                           est_free_estradiol_fc_se = "Standard error: Imputed free serum estradiol.",
                           est_free_estradiol_fc_low = "Lower 95% CI: Imputed free serum estradiol.",
                           est_free_estradiol_fc_high = "Upper 95% CI: Imputed free serum estradiol.",
                           est_progesterone_fc = paste("Imputed serum progesterone level based on the BioCycle study, loo-R =", loo_ci(mod_p_fc)),
                           est_progesterone_fc_se = "Standard error: Imputed serum progesterone",
                           est_progesterone_fc_low = "Lower 95% CI: Imputed serum progesterone",
                           est_progesterone_fc_high = "Upper 95% CI: Imputed serum progesterone",
                           est_testosterone_fc = paste("Imputed serum testosterone level based on the BioCycle study, loo-R =", loo_ci(mod_t_fc)),
                           est_testosterone_fc_se = "Standard error: Imputed serum testosterone",
                           est_testosterone_fc_low = "Lower 95% CI: Imputed serum testosterone",
                           est_testosterone_fc_high = "Upper 95% CI: Imputed serum testosterone",
                           est_free_testosterone_fc = paste("Imputed free serum testosterone level based on the BioCycle study, loo-R =", loo_ci(mod_ft_fc)),
                           est_free_testosterone_fc_se = "Standard error: Imputed free serum testosterone",
                           est_free_testosterone_fc_low = "Lower 95% CI: Imputed free serum testosterone",
                           est_free_testosterone_fc_high = "Upper 95% CI: Imputed free serum testosterone",
                           est_pbfw_fc = paste("Imputed probability of being in the fertile window based on the BioCycle study (LH test), loo-R =", loo_ci(mod_pbfw_fc)),
                           est_pbfw_fc_se = "Standard error: Imputed probability of being in the fertile window",
                           est_pbfw_fc_low = "Lower 95% CI: Imputed probability of being in the fertile window",
                           est_pbfw_fc_high = "Upper 95% CI: Imputed probability of being in the fertile window"
                           )

var_label(bc_days) <- list(bc_day = "Backward-counted cycle day. Counted from next observed menstrual onset (0), beginning with -1.",
                           prc_stirn_bc = "Probability of being in the fertile window determined according to the backward-counted cycle days, probabilities from Gangestad et al. 2016 based on Stirnemann et al. 2013",
                           prc_wcx_bc = "Probability of being in the fertile window determined according to the backward-counted cycle days, probabilities from Gangestad et al. 2016 based on Wilcox et al. 2001",
                           est_estradiol_bc = paste("Imputed serum estradiol level based on the BioCycle study, loo-R =", loo_ci(mod_e_bc)),
                           est_estradiol_bc_se = "Standard error: Imputed serum estradiol.",
                           est_estradiol_bc_low = "Lower 95% CI: Imputed serum estradiol.",
                           est_estradiol_bc_high = "Upper 95% CI: Imputed serum estradiol.",
                           est_free_estradiol_bc = paste("Imputed free serum estradiol level based on the BioCycle study, loo-R =", loo_ci(mod_fe_bc)),
                           est_free_estradiol_bc_se = "Standard error: Imputed free serum estradiol.",
                           est_free_estradiol_bc_low = "Lower 95% CI: Imputed free serum estradiol.",
                           est_free_estradiol_bc_high = "Upper 95% CI: Imputed free serum estradiol.",
                           est_progesterone_bc = paste("Imputed serum progesterone level based on the BioCycle study, loo-R =", loo_ci(mod_p_bc)),
                           est_progesterone_bc_se = "Standard error: Imputed serum progesterone",
                           est_progesterone_bc_low = "Lower 95% CI: Imputed serum progesterone",
                           est_progesterone_bc_high = "Upper 95% CI: Imputed serum progesterone",
                           est_testosterone_bc = paste("Imputed serum testosterone level based on the BioCycle study, loo-R =", loo_ci(mod_t_bc)),
                           est_testosterone_bc_se = "Standard error: Imputed serum testosterone",
                           est_testosterone_bc_low = "Lower 95% CI: Imputed serum testosterone",
                           est_testosterone_bc_high = "Upper 95% CI: Imputed serum testosterone",
                           est_free_testosterone_bc = paste("Imputed free serum testosterone level based on the BioCycle study, loo-R =", loo_ci(mod_ft_bc)),
                           est_free_testosterone_bc_se = "Standard error: Imputed free serum testosterone",
                           est_free_testosterone_bc_low = "Lower 95% CI: Imputed free serum testosterone",
                           est_free_testosterone_bc_high = "Upper 95% CI: Imputed free serum testosterone",
                           est_pbfw_bc = paste("Imputed probability of being in the fertile window based on the BioCycle study (LH test), loo-R =", loo_ci(mod_pbfw_bc)),
                           est_pbfw_bc_se = "Standard error: Imputed probability of being in the fertile window",
                           est_pbfw_bc_low = "Lower 95% CI: Imputed probability of being in the fertile window",
                           est_pbfw_bc_high = "Upper 95% CI: Imputed probability of being in the fertile window"
                           )

var_label(lh_days) <- list(lh_day = "Day relative to LH surge",
                           est_estradiol_lh = paste("Imputed serum estradiol level based on the BioCycle study, loo-R =", loo_ci(mod_e_lh)),
                           est_estradiol_lh_se = "Standard error: Imputed serum estradiol.",
                           est_estradiol_lh_low = "Lower 95% CI: Imputed serum estradiol.",
                           est_estradiol_lh_high = "Upper 95% CI: Imputed serum estradiol.",
                           est_free_estradiol_lh = paste("Imputed free serum estradiol level based on the BioCycle study, loo-R =", loo_ci(mod_fe_lh)),
                           est_free_estradiol_lh_se = "Standard error: Imputed free serum estradiol.",
                           est_free_estradiol_lh_low = "Lower 95% CI: Imputed free serum estradiol.",
                           est_free_estradiol_lh_high = "Upper 95% CI: Imputed free serum estradiol.",
                           est_progesterone_lh = paste("Imputed serum progesterone level based on the BioCycle study, loo-R =", loo_ci(mod_p_lh)),
                           est_progesterone_lh_se = "Standard error: Imputed serum progesterone",
                           est_progesterone_lh_low = "Lower 95% CI: Imputed serum progesterone",
                           est_progesterone_lh_high = "Upper 95% CI: Imputed serum progesterone",
                           est_testosterone_lh = paste("Imputed serum testosterone level based on the BioCycle study, loo-R =", loo_ci(mod_t_lh)),
                           est_testosterone_lh_se = "Standard error: Imputed serum testosterone",
                           est_testosterone_lh_low = "Lower 95% CI: Imputed serum testosterone",
                           est_testosterone_lh_high = "Upper 95% CI: Imputed serum testosterone",
                           est_free_testosterone_lh = paste("Imputed free serum testosterone level based on the BioCycle study, loo-R =", loo_ci(mod_ft_lh)),
                           est_free_testosterone_lh_se = "Standard error: Imputed free serum testosterone",
                           est_free_testosterone_lh_low = "Lower 95% CI: Imputed free serum testosterone",
                           est_free_testosterone_lh_high = "Upper 95% CI: Imputed free serum testosterone",
                           conception_risk_lh = "Risk of conception according to Stern, Kordsmeyer, & Penke 2021 Table 1 weighted averages.",
                           fertile_lh = "Probability of being in the fertile window according to Stern, Kordsmeyer, & Penke 2021 Table 1, taking highest conception risk as the denominator and brought to same mean as Stirnemann numbers."
                           )
```

## Save merge files
```{r}
rio::export(fc_days, "merge_files/fc_days.rds")
rio::export(bc_days, "merge_files/bc_days.rds")
rio::export(lh_days, "merge_files/lh_days.rds")
rio::export(fc_days, "merge_files/fc_days.sav")
rio::export(bc_days, "merge_files/bc_days.sav")
rio::export(lh_days, "merge_files/lh_days.sav")
rio::export(fc_days, "merge_files/fc_days.tsv")
rio::export(bc_days, "merge_files/bc_days.tsv")
rio::export(lh_days, "merge_files/lh_days.tsv")
```


# Codebooks
## Forward-counted
```{r}
codebook::compact_codebook(fc_days)
```

## Backward-counted
```{r}
codebook::compact_codebook(bc_days)
```

## Relative to LH surge
```{r}
codebook::compact_codebook(lh_days)
```

