---
title: "Multivariate Imputation of Hormones"
author: "Ruben Arslan"
date: "1/30/2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
options(width = 8000)
```

Here, I rerun the models documented in [Arslan et al. (2023)](https://dx.doi.org/10.1016/j.psyneuen.2022.105994). Not within spitting distance: Salivary immunoassays of estradiol have subpar validity for predicting cycle phase. 
The changes are:
- switched models to be multivariate with residual correlations between outcomes.
- dropped one outcome (fertile_lh) which was a side interest and led to many rows being dropped because of missingness
- added an outcome (log LH, luteinising hormone). It wasn't part of the original set and since I needed it for a different project, I added it here.

The imputed values changed only very slightly. I've documented the mistake of not using mutually adjusted multivariate predictors [here](https://rubenarslan.github.io/mistakes.html).

```{r}
library(tidyverse)
library(codebook)
library(brms)
library(rio)

plot_curve <- function(model, method = "fitted", resp = NULL) {
  cond <- tibble(.rows = 1)
  cond[[names(model$data)[2]]] <- "none"
  if (!is.null(resp)) {
    plot(conditional_effects(model, method = method,
                  conditions = cond, 
                  spaghetti = TRUE, nsamples = 100, resp = resp), 
                points = TRUE, 
                point_args = list(alpha = 0.05),
                spaghetti_args = list(alpha = 0.01, colour = '#FF000007'), 
                line_args = list(size = 0), 
                plot = FALSE)[[1]]
  } else {
    plot(conditional_effects(model, method = method,
                  conditions = cond, 
                  spaghetti = TRUE, nsamples = 100), 
                points = TRUE, 
                point_args = list(alpha = 0.05),
                spaghetti_args = list(alpha = 0.01, colour = '#FF000007'), 
                line_args = list(size = 0), 
                plot = FALSE)[[1]]
  }
}

rmse <- function(error) { sqrt(mean(error^2, na.rm = TRUE)) }
rmse_brms <- function(model) {
  resid <- residuals(model, summary = F)
  rmse_samples <- apply(resid, 1, rmse)
  rmse_ci <- rstantools::posterior_interval(as.matrix(rmse_samples), prob = 0.95)
  sprintf("%.2f [%.2f;%.2f]", mean(rmse_samples), rmse_ci[1], rmse_ci[2]) 
}
mae <- function(error) { mean(abs(error), na.rm = TRUE) }
mae_brms <- function(model) {
  resid <- residuals(model, summary = F)
  mae_samples <- apply(resid, 1, mae)
  mae_ci <- rstantools::posterior_interval(as.matrix(mae_samples), prob = 0.95)
  sprintf("%.2f [%.2f;%.2f]", mean(mae_samples), mae_ci[1], mae_ci[2]) 
}

loo_ci <- function(model) {
  loo_R <- loo_R2(model, re_formula = NA)
  loo_R <- sqrt(loo_R)
  sprintf("%.2f [%.2f;%.2f]", loo_R[,"Estimate"], loo_R[,"Q2.5"], loo_R[,"Q97.5"]) 
}
```

```{r}
biocycle <- readRDS("biocycle.rds")
options(mc.cores = parallel::detectCores(), brms.backend = "cmdstanr", brms.file_refit = "on_change")
rstan::rstan_options(auto_write = TRUE)

biocycle <- biocycle %>% filter(between(cycle_length, 20, 35))
biocycle$logE_P <- log(biocycle$E_P)
biocycle$logE <- log(biocycle$total_estradiol)
biocycle$logP <- log(biocycle$progesterone)
biocycle$logT <- log(biocycle$Testo)
biocycle$logFE <- log(biocycle$estradiol)
biocycle$logFT <- log(biocycle$`Free T (ng/dL)`)
biocycle$logLH <- log(biocycle$LH)
biocycle$fertile_lh <- NULL
```

## Days
```{r}
bc_days <- data.frame(
	bc_day = c(-28:-1, -29:-40),
	prc_stirn_bc = c(.01, .01, .02, .03, .05, .09, .16, .27, .38, .48, .56, .58, .55, .48, .38, .28, .20, .14, .10, .07, .06, .04, .03, .02, .01, .01, .01, .01, rep(.01, times = 12)),
# 	                rep(.01, times = 70)), # gangestad uses .01 here, but I think such cases are better thrown than kept, since we might simply have missed a mens
	prc_wcx_bc = c(.000, .000, .001, .002, .004, .009, .018, .032, .050, .069, .085, .094, .093, .085, .073, .059, .047, .036, .028, .021, .016, .013, .010, .008, .007, .006, .005, .005, rep(.005, times = 12))
) %>% arrange(bc_day)

fc_days <- data.frame(
	fc_day = c(0:39),
	prc_stirn_fc = c(.01, .01, .02, .03, .05, .09, .16, .27, .38, .48, .56, .58, .55, .48, .38, .28, .20, .14, .10, .07, .06, .04, .03, .02, .01, .01, .01, .01, rep(.01, times = 12)),
# 	                rep(.01, times = 70)), # gangestad uses .01 here, but I think such cases are better thrown than kept, since we might simply have missed a mens
	prc_wcx_fc = c(.000, .000, .001, .002, .004, .009, .018, .032, .050, .069, .085, .094, .093, .085, .073, .059, .047, .036, .028, .021, .016, .013, .010, .008, .007, .006, .005, .005, rep(.005, times = 12))
)

lh_days <- tibble(
    conception_risk_lh = c(rep(0,8), 0.00, 0.01, 0.02, 0.06, 0.16, 0.20, 0.25, 0.24, 0.10, 0.02, 0.02, rep(0,12)),
    lh_day = -15:15
  ) %>% 
    mutate(fertile_lh = conception_risk_lh/max(conception_risk_lh)*0.25/0.31)
biocycle <- biocycle %>% left_join(lh_days %>% select(lh_day, fertile_lh), by = "lh_day")
```

## Multivariate Imputation

### Backward counting

```{r}
bf_e_bc <- bf(logE ~ s(bc_day) + (1 | id) + (1|id:cycle))
bf_fe_bc <- bf(logFE ~ s(bc_day) + (1 | id) + (1|id:cycle))
bf_p_bc <- bf(logP ~ s(bc_day) + (1 | id) + (1|id:cycle))
bf_t_bc <- bf(logT ~ s(bc_day) + (1 | id) + (1|id:cycle))
bf_ft_bc <- bf(logFT ~ s(bc_day) + (1 | id) + (1|id:cycle))
bf_lh_bc <- bf(logLH ~ s(bc_day) + (1 | id) + (1|id:cycle))

mod_mv_bc <- brm(
  bf_e_bc + bf_fe_bc + bf_p_bc + bf_t_bc + bf_ft_bc + bf_lh_bc + set_rescor(TRUE),
  data = biocycle,
  control = list(adapt_delta = 0.95, max_treedepth = 20),
  file = 'models/mod_mv_bc'
)
mod_mv_bc

# Predict
preds_bc <- predict(mod_mv_bc, newdata = bc_days, re_formula = NA)

# Assign to bc_days using named list elements for safety
bc_days$mv_est_estradiol_bc <- round(preds_bc[, "Estimate", "logE"], 2)
bc_days$mv_est_free_estradiol_bc <- round(preds_bc[, "Estimate", "logFE"], 2)
bc_days$mv_est_progesterone_bc <- round(preds_bc[, "Estimate", "logP"], 2)
bc_days$mv_est_testosterone_bc <- round(preds_bc[, "Estimate", "logT"], 2)
bc_days$mv_est_free_testosterone_bc <- round(preds_bc[, "Estimate", "logFT"], 2)
bc_days$mv_est_luteinising_bc <- round(preds_bc[, "Estimate", "logLH"], 2)
rio::export(bc_days ,"merge_files/bc_days_mv.rds")
rio::export(bc_days ,"merge_files/bc_days_mv.tsv")
rio::export(bc_days ,"merge_files/bc_days_mv.sav")
```

### Forward counting

```{r}
bf_e_fc <- bf(logE ~ s(fc_day) + (1 | id) + (1|id:cycle))
bf_fe_fc <- bf(logFE ~ s(fc_day) + (1 | id) + (1|id:cycle))
bf_p_fc <- bf(logP ~ s(fc_day) + (1 | id) + (1|id:cycle))
bf_t_fc <- bf(logT ~ s(fc_day) + (1 | id) + (1|id:cycle))
bf_ft_fc <- bf(logFT ~ s(fc_day) + (1 | id) + (1|id:cycle))
bf_lh_fc <- bf(logLH ~ s(fc_day) + (1 | id) + (1|id:cycle))

mod_mv_fc <- brm(
  bf_e_fc + bf_fe_fc + bf_p_fc + bf_t_fc + bf_ft_fc + bf_lh_fc + set_rescor(TRUE),
  data = biocycle,
  control = list(adapt_delta = 0.95, max_treedepth = 20),
  file = 'models/mod_mv_fc'
)
mod_mv_fc

preds_fc <- predict(mod_mv_fc, newdata = fc_days, re_formula = NA)

fc_days$mv_est_estradiol_fc <- round(preds_fc[, "Estimate", "logE"], 2)
fc_days$mv_est_free_estradiol_fc <- round(preds_fc[, "Estimate", "logFE"], 2)
fc_days$mv_est_progesterone_fc <- round(preds_fc[, "Estimate", "logP"], 2)
fc_days$mv_est_testosterone_fc <- round(preds_fc[, "Estimate", "logT"], 2)
fc_days$mv_est_free_testosterone_fc <- round(preds_fc[, "Estimate", "logFT"], 2)
fc_days$mv_est_luteinising_fc <- round(preds_fc[, "Estimate", "logLH"], 2)
rio::export(fc_days ,"merge_files/fc_days_mv.rds")
rio::export(fc_days ,"merge_files/fc_days_mv.tsv")
rio::export(fc_days ,"merge_files/fc_days_mv.sav")
```

### Relative to LH surge

```{r}
# Filter data
biocycle_lh <- biocycle %>% filter(between(lh_day, -20, 14))

bf_e_lh <- bf(logE ~ s(lh_day) + (1 | id) + (1|id:cycle))
bf_fe_lh <- bf(logFE ~ s(lh_day) + (1 | id) + (1|id:cycle))
bf_p_lh <- bf(logP ~ s(lh_day) + (1 | id) + (1|id:cycle))
bf_t_lh <- bf(logT ~ s(lh_day) + (1 | id) + (1|id:cycle))
bf_ft_lh <- bf(logFT ~ s(lh_day) + (1 | id) + (1|id:cycle))
bf_lh_lh <- bf(logLH ~ s(lh_day) + (1 | id) + (1|id:cycle))

# PBFW not modeled in LH model

mod_mv_lh <- brm(
  bf_e_lh + bf_fe_lh + bf_p_lh + bf_t_lh + bf_ft_lh + bf_lh_lh +  set_rescor(TRUE),
  data = biocycle_lh,
  control = list(adapt_delta = 0.95, max_treedepth = 20),
  file = 'models/mod_mv_lh'
)
mod_mv_lh

preds_lh <- predict(mod_mv_lh, newdata = lh_days, re_formula = NA)

lh_days$mv_est_estradiol_lh <- round(preds_lh[, "Estimate", "logE"], 2)
lh_days$mv_est_free_estradiol_lh <- round(preds_lh[, "Estimate", "logFE"], 2)
lh_days$mv_est_progesterone_lh <- round(preds_lh[, "Estimate", "logP"], 2)
lh_days$mv_est_testosterone_lh <- round(preds_lh[, "Estimate", "logT"], 2)
lh_days$mv_est_luteinising_lh <- round(preds_lh[, "Estimate", "logLH"], 2)
lh_days$mv_est_free_testosterone_lh <- round(preds_lh[, "Estimate", "logFT"], 2)
rio::export(lh_days ,"merge_files/lh_days_mv.rds")
rio::export(lh_days ,"merge_files/lh_days_mv.tsv")
rio::export(lh_days ,"merge_files/lh_days_mv.sav")
```

## Comparison

```{r}
# Load old files
bc_days_old <- rio::import("merge_files/bc_days.rds")
fc_days_old <- rio::import("merge_files/fc_days.rds")
lh_days_old <- rio::import("merge_files/lh_days.rds")

# Join and correlate
# BC
bc_joined <- left_join(bc_days, bc_days_old, by = "bc_day", suffix = c("_new", "_old"))

cor_bc <- bc_joined %>% 
  summarise(
    cor_e = cor(mv_est_estradiol_bc, est_estradiol_bc, use = "complete.obs"),
    cor_fe = cor(mv_est_free_estradiol_bc, est_free_estradiol_bc, use = "complete.obs"),
    cor_p = cor(mv_est_progesterone_bc, est_progesterone_bc, use = "complete.obs"),
    cor_t = cor(mv_est_testosterone_bc, est_testosterone_bc, use = "complete.obs"),
    cor_ft = cor(mv_est_free_testosterone_bc, est_free_testosterone_bc, use = "complete.obs")
  )
print("Correlations for BC models:")
print(cor_bc)

# FC
fc_joined <- left_join(fc_days, fc_days_old, by = "fc_day", suffix = c("_new", "_old"))

cor_fc <- fc_joined %>% 
  summarise(
    cor_e = cor(mv_est_estradiol_fc, est_estradiol_fc, use = "complete.obs"),
    cor_fe = cor(mv_est_free_estradiol_fc, est_free_estradiol_fc, use = "complete.obs"),
    cor_p = cor(mv_est_progesterone_fc, est_progesterone_fc, use = "complete.obs"),
    cor_t = cor(mv_est_testosterone_fc, est_testosterone_fc, use = "complete.obs"),
    cor_ft = cor(mv_est_free_testosterone_fc, est_free_testosterone_fc, use = "complete.obs")
  )
print("Correlations for FC models:")
print(cor_fc)

# LH
lh_joined <- left_join(lh_days, lh_days_old, by = "lh_day", suffix = c("_new", "_old"))

cor_lh <- lh_joined %>% 
  summarise(
    cor_e = cor(mv_est_estradiol_lh, est_estradiol_lh, use = "complete.obs"),
    cor_fe = cor(mv_est_free_estradiol_lh, est_free_estradiol_lh, use = "complete.obs"),
    cor_p = cor(mv_est_progesterone_lh, est_progesterone_lh, use = "complete.obs"),
    cor_t = cor(mv_est_testosterone_lh, est_testosterone_lh, use = "complete.obs"),
    cor_ft = cor(mv_est_free_testosterone_lh, est_free_testosterone_lh, use = "complete.obs")
  )
print("Correlations for LH models:")
print(cor_lh)
```

## Bias Investigation: Inter-hormone Correlations

This section checks if the univariate imputation (Old) underestimated the correlation *between* different hormones compared to the multivariate imputation (New).

```{r}
# Function to get lower triangle of correlation matrix
get_cor_summary <- function(df, pattern) {
  # Select columns matching pattern
  cols <- df %>% select(matches(pattern))
  # Calculate correlation
  cormat <- cor(cols, use = "complete.obs")
  # Return readable format
  return(cormat)
}

print("--- Backward Counting: Inter-hormone correlations ---")

# Old Correlations (Univariate)
print("OLD (Univariate) Inter-hormone Correlations:")
old_vars_bc <- c("est_estradiol_bc", "est_progesterone_bc", "est_testosterone_bc", "est_free_estradiol_bc")
print(cor(bc_joined[, old_vars_bc], use = "complete.obs"))

# New Correlations (Multivariate)
print("NEW (Multivariate) Inter-hormone Correlations:")
new_vars_bc <- c("mv_est_estradiol_bc", "mv_est_progesterone_bc", "mv_est_testosterone_bc", "mv_est_free_estradiol_bc")
print(cor(bc_joined[, new_vars_bc], use = "complete.obs"))

print("--- Difference (New - Old) ---")
# Calculate difference matrix (matching names first)
c_old <- cor(bc_joined[, old_vars_bc], use = "complete.obs")
c_new <- cor(bc_joined[, new_vars_bc], use = "complete.obs")
dimnames(c_new) <- dimnames(c_old) # Ensure names match for subtraction
print(round(c_new - c_old, 3))
```

## Summary of Findings

Based on the comparison between the old univariate imputations and the new multivariate imputations (with residual correlations modeled):

1.  **Backward Counting (BC):**
    *   **Negligible Bias:** The correlations between old and new estimates are >0.98.
    *   **Conclusion:** The univariate method performed nearly identically to the multivariate method.
    
2.  **Forward Counting (FC):**
    *   **Negligible Bias:** The correlations between old and new estimates are >0.97.
    *   **Conclusion:** The univariate method performed nearly identically to the multivariate method.

3.  **LH-centered Method (LH):**
    *   **No Bias:** The correlation between old and new estimates is ~1.00.
    *   **Conclusion:** The LH-centered method is robust. The cycle phases are so well-aligned by the LH surge that the mean curve explains almost all variance, and multivariate information adds virtually nothing.

